---
title: "Portfolio Analysis - Core Fund - April 2025"
author: "Financial Analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: cosmo
    highlight: tango
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                     fig.width = 10, fig.height = 6)
```

# Introduction

This document presents a comprehensive analysis of a revised proposed portfolio - our **Core Fund** - intended to protect against currency debasement risks, while maintaining sufficient exposure to the broad equity market to capture upside returns. The AJ Bell platform will be used to manage this portfolio. Around 90% of our assets will be managed through this fund. 

Alongside this, we will run a much smaller high risk fund - our **Venture Fund** - which will focus on investing in individual companies and cryptoassets. This will account for around 10% of assets under management. 

# The Core Fund

The individual funds to be included in the *Core Fund* portfolio are as follows:

* **iShares Physical Gold ETC**: A [fund](https://www.ishares.com/uk/individual/en/products/258441/ishares-physical-gold-etc-fund?switchLocale=y&siteEntryPassthrough=true) which closely tracks the London Bullion Market Association (LBMA) Gold Price - the most widely used benchmark for daily gold prices. It's base currency is USD. The fund was launched on the 8th April 2011. It's ticker is SGLN (SGLN.L on Yahoo Finance). Indicative Fund Share - 15%.
* **iShares Core MSCI World UCITS ETF**: The broad equity component of our portfolio. This fund provides exposure to a wide range of equities across 23 developed countries. As of March 2025, US equities accounted for 71% of the portfolio. The investor documentation can be found [here](https://www.ishares.com/uk/individual/en/products/287737/ishares-core-msci-world-ucits-etf). The fund was launched on the 22nd May 2017. It is GBP hedged. It's ticker is IWDG (IWDG.L on Yahoo Finance). Indicative Fund Share - 40%. 
* **iShares Edge MSCI World Value Factor UCITS ETF**: Provides exposure to a subset of MSCI world stocks that are undervalued relative to their fundamentals. As of April 2025, 38% of the fund is invested in US stocks, 21% in Japanese stocks, and around 10% in UK stocks. There is not a particularly strong skew towards defensive sectors - Information Technology remains the number one sector at 23%. The ticker is IWVL (IWVL.L on Yahoo Finance). It is USD denominated, so comes with currency risk. The investor documentation can be found [here](https://www.ishares.com/uk/individual/en/products/270048/ishares-msci-world-value-factor-ucits-etf). Indicative Fund Share - 10%.
* **iShares Edge MSCI World Minimum Volatility UCITS ETF**: Shares in a set of companies from developed countries that have lower volatility characteristics. There is subsequently a skew towards consumer staples, industrials and utilities, though cyclical sectors such as information technology retain a high share. The fund was launched on 30th November 2012. I high percentage (around 70% of AUM is invested in US-listed companies). The investor documentation can be found [here](https://www.ishares.com/uk/individual/en/products/251382/ishares-msci-world-minimum-volatility-ucits-etf). The fund is USD denominated, and therefore comes with currency risk. Indicative Fund Share - 10%. Yahoo Finance ticker MVOL.L. 
* **iShares MSCI Emerging Markets ex China ETF**: USD denominated, and therefore carries currency risk. The fund looks to gain exposure to large and mid cap companies in emerging markets (excluding China, where as UK citizens we are worried about property rights). It has large holdings in Information Technology (over 30%), including a significant TSMC holding (14%). Launched in July 2017. The investor documentation can be found [here](https://www.ishares.com/us/products/288504/ishares-msci-emerging-markets-ex-china-etf). The Yahoo Finance ticker is EMXC. Indicative Fund Share - 5%. 
* **Future of Defence UCITS ETF**. The stock ticker is NATP.L on Yahoo Finance. It has a relatively recent inception date in 2023. It is GBP denominated. The fund has broad cross-country exposure - with the US the largest share. It provides exposure to companies generating revenues from NATO defense and cyber-defense spending. We will use Invesco Aerospace and Defence ETF (Yahoo Finance ticker PPA) as an imperfect proxy with a longer backrun for portfolio analysis. Indicative allocation - 10%. Investment documentation [here](https://hanetf.com/fund/nato-future-of-defence-etf/#tab-id-1). 
* **Invesco Bloomberg Commodity UCITS ETF**: Yahoo Finance ticker - CMOP.L. Comprised of futures contracts on physical commodities. 24 commodities are eligible for inclusion across energy, grains, industrial metals and livestock. The fund has an inception date in 2018. Indicative Fund Share - 10%. Investor documentation [here](https://www.hl.co.uk/shares/shares-search-results/i/invesco-bloomberg-commodity-ucits-etf-gbp). 

*Potential future candidate fund*

* **iShares Global Inflation Linked Govt Bond UCITS ETF**: Yahoo Finance ticker IGIL.L. Provides diversified exposure to investment grade inflation-linked government bonds. The largest shares are US and UK inflation-linked government bonds. Fund documentation [here](https://www.ishares.com/uk/individual/en/products/251746/ishares-global-inflation-linked-government-bond-ucits-etf). The fund shares are USD denominated, and so come with currency risk. The inception date in August 2008. Indicative Fund Share - 0% (until returns are adequately understood by us). 

```{r }

# Load required libraries
library(quantmod)
library(PerformanceAnalytics)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(reshape2)
library(scales)
library(gridExtra)
library(zoo)  # For rolling calculations
library(magrittr)
library(xts)
library(knitr)
library(kableExtra)
library(RColorBrewer)
library(corrplot)

# Set theme for consistent visualization
theme_set(theme_minimal() + 
          theme(plot.title = element_text(face = "bold"),
                plot.subtitle = element_text(face = "italic"),
                legend.position = "bottom"))

```

## Visualisation of Portfolio Composition

```{r }

# Define the portfolio allocation
ticker_names <- c("Physical Gold", "Core MSCI World", "MSCI World Value Factor", "MSCI World Minimum Volatility",
                  "MSCI Emerging Markets ex China", "Future of Defence", "Bloomberg Commodity")
tickers <- c("SGLN.L", "IWDG.L", "IWVL.L", "MVOL.L", "EMXC", "NATP.L", "CMOP.L")

portfolio_allocation <- tibble(
  Fund = ticker_names,
  Ticker = tickers,
  Allocation = c(0.15, 0.40, 0.1, 0.1, 0.05, 0.1, 0.1) # Update these in future months as appropriate. 
)

# Create a pie chart of allocations
ggplot(portfolio_allocation, aes(x = "", y = Allocation, fill = Fund)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Core Fund Portfolio Allocation",
       subtitle = "April 2025 - Initial Investment Proposal Allocation",
       fill = "Investment Fund") +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = percent_format()) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))
```

# Theory behind The Core Fund Composition

[**To be completed**]

# Important Fund Considerations

## Using Simple or Log Returns for Portfolio Analysis

In this analysis we will work with **log returns** for all statistical computations, and convert back to simple (arithmetic) returns only when reporting or interpreting performance.

### Definitions

Let \(P_t\) be the price (or index level) at the end of month \(t\). We define:

- **Log return**  
  \[
    r_t = \ln\!\bigl(\tfrac{P_t}{P_{t-1}}\bigr)
  \]
- **Re‐conversion to simple return**  
  \[
    R_t = \exp(r_t) - 1
  \]

### Justification for Log Returns

1. **Time additivity**  
   
   \[
     r_{1\to T} \;=\; \sum_{t=2}^T r_t
     \quad\Longleftrightarrow\quad
     R_{1\to T} \;=\; \exp\!\Bigl(\sum_{t=2}^T r_t\Bigr)\;-\;1
   \]
   
   This means multi‐period compounding reduces to a simple sum of \(r_t\).

2. **Statistical properties** 

   - Log returns are often **closer to Normally distributed** (smaller skewness, lower excess kurtosis), which makes:
     - **Estimation** of means and variances more reliable.
     - **Modeling** (e.g. GARCH, factor models) directly applicable under Gaussian assumptions.

3. **Portfolio aggregation**  
   
   A portfolio log return is simply the weighted sum of individual asset log returns:
   \[
     r_{p,t} = \sum_{i=1}^N w_i\,r_{i,t},
     \quad \sum_i w_i = 1
   \]
   
   No need for iterative compounding each period.

4. **Large‐move symmetry**  
   
   For large price swings, \(\ln\)-returns treat gains and losses symmetrically (no upper‐bound bias).

# Analysis of Core Fund Components

This section... [**To be completed**]

## Fund Prices Across Time

### Choice to use Adjusted Closing Prices

Adjusted closing prices incorporate dividends, rights issues, stock splits and other corporate actions into a single, continuous price series, ensuring that return calculations truly reflect the value changes experienced by investors. Without these adjustments, raw closing prices would show artificial jumps or drops on ex‑dividend or split dates, distorting measures of performance, volatility and correlation. By using adjusted prices, all dividends and distributions are effectively reinvested in the series, yielding total‑return log‑returns that are comparable across time and across securities. This consistency is essential for accurate performance attribution, portfolio optimisation and risk analysis, and it guarantees that backtests and historical studies align with the real economic experience of holding the assets.

### Visualisation of Price Time Series

In our portfolio analysis, we will use monthly returns. 

```{r message = FALSE, fig.height=18, fig.width=10}

# Assign start and end dates
end_date <- Sys.Date()
start_date <- as.Date("2020-01-01")

# Create empty list to store price data
price_data <- list()

# Download data for each ticker
for (i in 1:length(tickers)) {
  ticker <- tickers[i]
  
 # Try to download data, with error handling
  tryCatch({
    getSymbols(ticker, src = "yahoo",
               # from = start_date,
               # to = end_date,
               auto.assign = FALSE) -> price_data[[ticker]]

  #   cat(paste("Successfully downloaded data for", ticker, "\n"))
  # }, error = function(e) {
  #   cat(paste("Error downloading data for", ticker, ":", e$message, "\n"))
  })
}

# Function to extract closing prices and filter to month-end
get_monthly_prices <- function(price_series, ticker) {
  # Extract adjusted closing prices
  closing_prices <- Ad(price_series)
  
  # Convert to data frame for easier manipulation
  df <- data.frame(Date = index(closing_prices),
                   Price = as.numeric(closing_prices),
                   stringsAsFactors = FALSE)
  
  # Convert Date column to Date format
  df$Date <- as.Date(df$Date)
  
  # Extract year and month
  df$Year <- year(df$Date)
  df$Month <- month(df$Date)
  
  # Group by year and month, then get the last trading day of each month
  monthly_data <- df %>%
    group_by(Year, Month) %>%
    filter(Date == max(Date)) %>%
    ungroup()
  
  # Clean up the data frame
  monthly_data <- monthly_data %>%
    select(Date, Price) %>%
    mutate(Ticker = ticker)
  
  return(monthly_data)
}

# Apply function to each ticker
monthly_prices <- list()
for (ticker in names(price_data)) {
  monthly_prices[[ticker]] <- get_monthly_prices(price_data[[ticker]], ticker)
}

# Combine all monthly prices into one data frame
all_monthly_prices <- bind_rows(monthly_prices)

# The final date of the trading month is not always consistent, so we need to align these.
all_monthly_prices %<>%
  mutate(Date = as.Date(format(Date, "%Y-%m-01")) + months(1) - 1)

# Create mapping for prettier names
name_mapping <- setNames(ticker_names, tickers)

# Add fund names
all_monthly_prices$Fund <- name_mapping[all_monthly_prices$Ticker]

# Check for NA's in the data
NA_values <- all_monthly_prices %>%
  dplyr:: filter(is.na(Price)) ## good, there are no NA values
rm(NA_values)

# Plot each fund's monthly price series
ggplot(all_monthly_prices, aes(x = Date, y = Price, color = Fund)) +
  geom_line(linewidth = 1) +
  facet_wrap(~Fund, scales = "free_y", ncol = 1) +
  labs(title = "Monthly Share Prices",
       subtitle = "End-of-month prices for each fund",
       x = "Date",
       y = "Price",
       color = "Fund") +
  scale_x_date(date_labels = "%Y", date_breaks = "3 years") +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = "lightgray"))

```

## Currency Risk 

A number of the funds in the portfolio are USD denominated - and currency risk needs to be accounted for before we can calculate true returns to us as GBP investors. 

### Total GBP Log Return

When we hold a USD‑denominated fund but measure performance in GBP, the **total log return** is the sum of the local (USD) log return and the FX log return:

$$
r_{t}^{\text{GBP}}
\;=\;
\underbrace{\ln\!\Bigl(\tfrac{P_{t}^{\text{USD}}}{P_{t-1}^{\text{USD}}}\Bigr)}_{r_{t}^{\text{USD}}}
\;+\;
\underbrace{\ln\!\Bigl(\tfrac{S_{t}}{S_{t-1}}\Bigr)}_{r_{t}^{\text{FX}}},
$$

where:

- \(P_{t}^{\text{USD}}\) is the fund price in USD at time \(t\),

- \(S_t\) is the GBP‑per‑USD spot rate at time \(t\).

### Implications for Risk and Return

As a UK investor holding a USD‑quoted fund, our sterling‑returns consist of both the fund’s performance in US dollars and the FX move between USD and GBP. If the dollar strengthens versus the pound, our GBP returns get a boost (and vice versa).

## Proxy for Future of Defence Fund

We use the Invesco Aerospace and Defence ETF. Comparing a USD fund to a GBP fund means that currency risk needs to be taken account of before correlation analysis. The correlation analysis shows that PPA is very much an imperfect proxy for NATP, but it'll give us an ok idea. 

```{r message = FALSE}
# 1. Download the two price series
getSymbols(c("PPA", "NATP.L"))
# rm(list = ls())
# 1b. Adjust for currency difference of PPA 
getSymbols("GBPUSD=X") # Download exchange rate
fx_gbp_per_usd <- 1/Ad(`GBPUSD=X`) # create new xts object for conversions
colnames(fx_gbp_per_usd) <- ("gbpusd")

# merge data 
PPA_GBP = merge(Ad(PPA), fx_gbp_per_usd, join = "left")

# fill any days where FX is missing by carrying the last available rate forward
PPA_GBP$gbpusd <- na.locf(PPA_GBP$gbpusd)

# 4. Compute your fund price in GBP
ppa_gbp <- PPA_GBP$PPA.Adjusted * PPA_GBP$gbpusd
colnames(ppa_gbp) <- "ppa_gbp"

# Now price_gbp is your USD‐fund prices translated into GBP on every fund date.


# 2. Extract Adjusted Close and merge into one xts object
proxy_data <- merge(
  ppa_gbp,
  Ad(NATP.L)
)

# Assign the names to the two funds
colnames(proxy_data) <- c("PPA_GBP", "NATP.L")

# 3. Compute daily log returns, and drop any leading NAs
log_ret_proxy <- na.omit(diff(log(proxy_data)))
# head(log_ret_proxy)

# 4. Correlation matrix of the two series
corr_matrix_proxy <- cor(log_ret_proxy)

# 5. Small correlation plot
corrplot(
  corr_matrix_proxy,
  method     = "number",   # show the correlation coefficient
  type       = "lower",    # only lower triangle
  tl.cex     = 0.8,        # label size
  number.cex = 1.2,        # coefficient size
  diag       = TRUE       # hide diagonal
)

```

## Monthly and Annual Returns of Each Fund

### Adjust returns of USD funds so that they are represented in GBP terms

This shows that there is a reasonable level of diversification in the fund, with our emerging market exposures, commodities and gold doing the heavy lifting in terms of providing significant diversification benefits. 

```{r message = FALSE}

# 1. Download the data series
ticker_names <- c("Physical Gold", "Core MSCI World", "MSCI World Value Factor", "MSCI World Minimum Volatility",
                  "MSCI Emerging Markets ex China", "Aerospace & Defence", "Bloomberg Commodity", "GBPUSD")
tickers <- c("SGLN.L", "IWDG.L", "IWVL.L", "MVOL.L", "EMXC", "PPA", "CMOP.L", "GBPUSD=X")
getSymbols(tickers)

# 2. Extract adjusted closes
gold_USD        <- Ad(SGLN.L)
mci_world     <- Ad(IWDG.L)
world_value_USD    <- Ad(IWVL.L)
min_vol_USD      <- Ad(MVOL.L)
em_mkts_USD   <- Ad(EMXC)
defence_USD   <- Ad(PPA)
commodities    <- Ad(CMOP.L)
fx_usd_per_gbp <- Ad(`GBPUSD=X`)        # USD per 1 GBP

# 3. Convert to GBP per USD, since USD funds are quoted in USD
fx_gbp_per_usd <- 1 / fx_usd_per_gbp

# all the names of the funds 
ad_prices <- c("gold_USD", "mci_world", "world_value_USD", "min_vol_USD",
               "em_mkts_USD", "defence_USD", "commodities", "fx_gbp_per_usd")

# put into a list
ad_prices_list <- mget(ad_prices)

# Take the last price of each month, ignoring NAs
monthly_list <- lapply(ad_prices_list, function(x) {
  apply.monthly(na.omit(x), last)
})

# head(monthly_list$gold_USD)

# helper to shift an xts index to the calendar month end
set_to_month_end <- function(x) {
  # x is an xts with one obs per month on last trading day
  new_idx <- as.Date(as.yearmon(index(x)), frac = 1)
  index(x) <- new_idx
  x
}

# apply to each series in your list
monthly_calendar_list <- lapply(monthly_list, set_to_month_end)

# head(monthly_calendar_list$gold_USD)

# merge all prices into a single xts object
merged_ad_prices <- do.call(merge, monthly_calendar_list)

# tail(merged_ad_prices)

# calculate log returns, at the moment in different currencies
unadjusted_log_returns <- diff(log(merged_ad_prices))

# tail(unadjusted_log_returns)

# set column names
colnames(unadjusted_log_returns) <- ad_prices

# select the names of the columns you want to adjust to gbp returns
to_adj <- c("gold_USD", "world_value_USD", "min_vol_USD", "em_mkts_USD", "defence_USD")

# create a new xts object to manipulate 
gbp_log_returns <- unadjusted_log_returns

# drop=TRUE pulls out a plain vector of length N
fx_vec <- gbp_log_returns[, "fx_gbp_per_usd", drop = TRUE]  

# now fx_vec is length N, so R will add it to each of the k columns:
gbp_log_returns[, to_adj] <- gbp_log_returns[, to_adj] + fx_vec

# set new column names 
colnames(gbp_log_returns) <- ad_prices <- c("gold_gbp", "mci_world", "world_value_gbp", "min_vol_gbp",
               "em_mkts_gbp", "defence_gbp", "commodities", "fx_gbp_per_usd")

# tail(gbp_log_returns)
# tail(unadjusted_log_returns)

# 1. Drop the FX column (either by position or name)
rets2 <- gbp_log_returns[, colnames(gbp_log_returns) != "fx_gbp_per_usd"]

# 2. Restrict to the overlapping period (i.e. only months with no NA across any series)
rets_common <- na.omit(rets2)

# 3. Plot the correlation matrix with chart.Correlation
chart.Correlation(
  rets_common,
  histogram   = TRUE,    # show distributions on the diagonal
  pch         = 19,      # solid circles for the scatterplots
  main        = "Correlation of Monthly GBP Log‑Returns"
)

# rm("gold_USD", "mci_world", "world_value_USD", "min_vol_USD",
#                "em_mkts_USD", "defence_USD", "commodities", "fx_gbp_per_usd")

# [1] "ret_a" "ret_b" "ret_c"

```

## Volatility of Each Fund

This chart shows the 12 month rolling standard deviations of the log returns of each individual fund component. 

```{r }

# 1. Compute 12‑month rolling SD for all columns at once
rolling_sd <- rollapply(
  data      = rets_common,
  width     = 12,
  FUN       = function(x) apply(x, 2, sd),  # on each 12×k block, compute sd of each column
  by.column = FALSE,                        # FUN expects the full matrix
  align     = "right",
  fill      = NA
)

# 2. (Optional) name the columns if they got lost
colnames(rolling_sd) <- colnames(rets_common)
rolling_sd %<>% na.omit()

# 3. Plot all on one chart
chart.TimeSeries(
  rolling_sd,
  main       = "12 Month Rolling Volatility",
  ylab       = "Std Dev of Monthly Returns",
  legend.loc = "topright",
  lwd        = 2
)


```

## Other Characteristics of Each Fund

### VaR and ES

### Skewness and Kurtosis

# Portfolio Returns

This section examines the performance of the portfolio overall. 

## Against S&P 500

```{r }

# 1. Get S&P 500 and compute monthly log‐returns
getSymbols("^GSPC", src="yahoo", auto.assign = TRUE)
spx_monthly <- to.monthly(Ad(GSPC), indexAt = "lastof", OHLC = FALSE) # Understand why this gives the last calendar day of month

# Set the S&P 500 in GBP terms
fx_gbp_per_usd <- 1/Ad(`GBPUSD=X`) # create new xts object for conversions
fx_gbp_per_usd <- to.monthly(fx_gbp_per_usd, indexAt = "lastof", OHLC = FALSE)

colnames(fx_gbp_per_usd) <- ("gbpusd")

# merge data 
spx_gbp = merge(spx_monthly, fx_gbp_per_usd, join = "left")

# fill any days where FX is missing by carrying the last available rate forward
spx_gbp$gbpusd <- na.locf(spx_gbp$gbpusd)
colnames(spx_gbp)

# 4. Compute your index price in GBP
spx_monthly <- spx_gbp$GSPC.Adjusted * spx_gbp$gbpusd

spx_log     <- na.omit(diff(log(spx_monthly)))
colnames(spx_log) <- "SP500"

# 2. Align SPX to your fund series (rets_common)
# common_idx     <- intersect(index(rets_common), index(spx_log))
# rets_common    <- rets_common[common_idx, ]
spx_log        <- spx_log[index(rets_common)]

# 3. Convert log‐returns → simple returns
rets_simple <- exp(rets_common) - 1
spx_simple  <- exp(spx_log)     - 1



# 4. Quarterly‐rebalanced portfolio simple returns
weights   <- c(0.15, 0.40, 0.10, 0.10, 0.05, 0.10, 0.10)
pf_simple <- Return.portfolio(
  R             = rets_simple,
  weights       = weights,
  rebalance_on  = "quarters",
  geometric     = TRUE
)

# (Optional) if you really need log portfolio returns:
# pf_log <- log(1 + pf_simple)

# 5. Build cumulative return indices (growth of 1)
pf_cum   <- cumprod(1 + pf_simple)
spx_cum  <- cumprod(1 + spx_simple)
comp_cum <- cumprod(1 + rets_simple)

# 6a. Plot: Portfolio vs. S&P 500
chart.TimeSeries(
  cbind(Portfolio = pf_cum, SP500 = spx_cum),
  legend.loc = "topleft",
  main       = "Cumulative Returns: Portfolio vs S&P 500",
  ylab       = "Growth of £1",
  lwd        = 2
)
```

## Against Portfolio Components

```{r }
# 6b. Plot: Portfolio vs. Components
chart.TimeSeries(
  cbind(Portfolio = pf_cum, comp_cum),
  legend.loc = "topleft",
  main       = "Cumulative Returns: Portfolio vs. Components",
  ylab       = "Growth of £1",
  lwd        = 2
)

```

## Standard deviation comparisons

### Against S&P 500

```{r }

# install.packages(c("quantmod","PerformanceAnalytics"))  # if needed

# 1. Re‑compute SPX simple returns
# getSymbols("^GSPC", src = "yahoo", auto.assign = TRUE)
# spx_monthly <- to.monthly(Ad(GSPC), indexAt = "lastof", OHLC = FALSE)
# spx_log     <- na.omit(diff(log(spx_monthly)))
# spx_simple  <- exp(spx_log) - 1
# colnames(spx_simple) <- "SP500"

# 2. Prepare fund simple returns
# funds_simple <- exp(rets_common) - 1 # rets simple
# 
# # 3. Quarterly‑rebalanced portfolio simple returns
# weights   <- c(0.15, 0.40, 0.10, 0.10, 0.05, 0.10, 0.10)
# pf_simple <- Return.portfolio(
#   R             = funds_simple,
#   weights       = weights,
#   rebalance_on  = "quarters",
#   geometric     = FALSE
# )

# 4a. Merge PF vs SPX
vol_spx <- merge(pf = pf_simple, SP500 = spx_simple, join = "inner")

# 4b. 12‑month rolling SD for PF & SPX
rolling_vol_spx <- rollapply(
  vol_spx,
  width     = 12,
  FUN       = function(x) apply(x, 2, sd),
  by.column = FALSE,
  align     = "right",
  fill      = NA
)
rolling_vol_spx %<>% na.omit()
# 5a. Plot Portfolio vs. S&P 500 rolling vol
chart.TimeSeries(
  rolling_vol_spx,
  main       = "12‑Month Rolling Volatility: Portfolio vs. S&P 500",
  ylab       = "Rolling SD of Monthly Returns",
  legend.loc = "topleft",
  lwd        = 2
)

```

### Against Portfolio Components

```{r }
# 4c. Merge PF vs Components
vol_comp <- merge(pf = pf_simple, rets_simple, join = "inner")

# 4d. 12‑month rolling SD for PF & each fund
rolling_vol_comp <- rollapply(
  vol_comp,
  width     = 12,
  FUN       = function(x) apply(x, 2, sd),
  by.column = FALSE,
  align     = "right",
  fill      = NA
)

rolling_vol_comp %<>% na.omit()
# 5b. Plot Portfolio vs. Components rolling vol
chart.TimeSeries(
  rolling_vol_comp,
  main       = "12‑Month Rolling Volatility: Portfolio vs. Funds",
  ylab       = "Rolling SD of Monthly Returns",
  legend.loc = "topleft",
  lwd        = 2
)

```

## Plotting annual returns

```{r fig.height = 18, fig.width = 10}

#── 2. combine portfolio + funds ────────────────────────────────────────────────
all_rets <- cbind(
  Portfolio = pf_simple,
  rets_simple
)

#── 3. annual returns 2018–2024 ─────────────────────────────────────────────────
ann_rets_xts <- apply.yearly(
  all_rets,
  function(x) apply(x, 2, function(col) prod(1 + col) - 1)
)
# keep only full years 2018–2024
ann_rets_xts <- ann_rets_xts["2018/2024"]

# convert to data.frame
ann_df <- data.frame(
  Year = as.integer(format(index(ann_rets_xts), "%Y")),
  coredata(ann_rets_xts),
  row.names = NULL,
  check.names = FALSE
)

#── 4. compute 2025 YTD ─────────────────────────────────────────────────────────
ytd_vals <- apply(
  all_rets["2025"],
  2,
  function(col) prod(1 + col) - 1
)
ytd_df <- data.frame(
  Year = 2025,
  t(ytd_vals),
  row.names = NULL,
  check.names = FALSE
)

#── 5. combine full years + YTD ─────────────────────────────────────────────────
combined_df <- bind_rows(ann_df, ytd_df)

#── 6. reshape for ggplot ───────────────────────────────────────────────────────
long_df <- combined_df %>%
  pivot_longer(
    cols = -Year,
    names_to  = "Asset",
    values_to = "Return"
  )

#── 7. faceted bar chart ─────────────────────────────────────────────────────────
p <- ggplot(long_df, aes(x = factor(Year), y = Return)) +
  geom_col(aes(fill = ifelse(Return < 0, "red4", "green4")), width = 0.7) +
  facet_wrap(~ Asset, scales = "free_y", ncol = 1) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  geom_hline(yintercept = 0, colour = "black", linetype = "dashed") +
  labs(
    title = "Annual and YTD Simple Returns: Portfolio & Funds",
    x     = "Year",
    y     = "Return"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    strip.text       = element_text(face = "bold"),
    axis.text.x      = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    plot.title       = element_text(face = "bold", hjust = 0.5),
    legend.position = "none"
  )

print(p)

```



```{r }

#── 8. kable table ───────────────────────────────────────────────────────────────
combined_df %>%
  mutate(across(-Year, ~ .x * 100)) %>%
  kable(
    format  = "html",
    digits  = 2,
    caption = "Annual (2018–2024) and 2025 YTD Simple Returns (%) – Portfolio & Funds"
  ) %>%
  kable_styling(full_width = FALSE, position = "center")


```

## Portfolio Rebalancing Decision




## Cumulative Returns 

[**Include the S&P 500 as a benchmark**]





